State of The Practice in Indirect Conflicts: Causes, Mitigation Strategies, and Future Directions

### 2014-03-16 iced-2013/paper.tex
ICED final commit.

% Paper title
% Can use linebreaks \\ within to get better formatting as desired
\title{State of The Practice in Indirect Conflicts: Work Flows, Tools, and Future Direction}

@@

As Software Configuration Management (SCM) has grown, the maturity and norm of parallel
development has become the standard development process instead of the exception. With this parallel development
comes the need for larger awareness among developers to have ``an understanding of the activities of others
which provides a context for one's own activities''~\cite{Dourish:1992:ACS}. This added awareness
mitigates some downsides of parallel development which include the cost of conflict prevention and resolution. However,
empirical evidence shows that these mitigated losses continue to appear quite frequently and can prove to be a significant
and time-consuming chore for developers~\cite{Perry:2001:PCL}.

@@

While indirect conflict tools have shown potential from studies of developers, some of the same problems continue
to arise throughout most, if not all tools. The most prevalent issue is that of false positives and information
overload, which causes tools to eventually be
ignored~\cite{Sarma:2007:TSA, Servant:2010:CPI}. A second primary issue is that of dependency identification and
tracking. Many different dependencies have been proposed and used in indirect conflict tools such as method
invocation~\cite{Trainer:2005:BGT}, and class signatures~\cite{Sarma:2007:TSA} with varying success, but the
identification of failure inducing changes, other than those which are already identifiable by other means such
as compilers, and unit tests, to these dependencies still remains an issue.
Some tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA} have also incorporated social factors 
such as Cataldo et al.'s~\cite{Cataldo:2006:ICR} notion of socio-technical
congruence.
However, the same aforementioned problems still arise. These mentioned tools have not been able to study their own 
problems as their evaluations have
been small scale, involving a limited number of industrial participants or student evaluators.

We want to understand why it is so hard to tackle indirect conflicts, specifically through tool based solutions.
To do so, step back and intend to obtain a broader view of indirect conflicts
with a large field study of practitioners. We investigate the root causes of information overload, false positives, and
scalability issues to better understand why indirect conflict tools fail to achieve
the success of other domain tools. We determine: general events which 
cause indirect conflicts, when they occur, if compounding conditions exist, mitigation strategies developers
use, and what practitioners want from new tools.
We ask:

@@

We interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory questionnaire of 78
developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. Our findings indicate that:
\begin{itemize}
	\item Indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding
	\item Developers tend to prefer to use detection and resolution processes or tools over those of prevention
	\item developers do not want awareness mechanisms which provide non actionable results
	\item There exists a gap in software evolution analytical tools from the reliance on static analysis resulting in missed
				context of indirect conflicts
\end{itemize}

@@

Many indirect conflict tools have been created, tested, and published. Sarma et al.~\cite{Sarma:2007:TSA} created Palantir,
which can both detect potential indirect conflicts, at the class signature level, and alert developers to these conflicts.
Palantir represented one of the first serious attempts at aiding developers towards indirect conflicts. Holmes et
al.~\cite{Holmes:2010:CAR} take it one step further with their tool YooHoo, by detecting fine grained source code changes,
such as method return type changes, and create a taxonomy for different types of changes and their proneness to cause
indirect conflicts. This tool and its taxonomy had a severely reduced false positive rate, however, the true positives
detected may already be detectable by current tools such as compilers and unit testing. The tool Ariadne~\cite{Trainer:2005:BGT}
creates an environment where developers can see how source code changes will affect other areas of a project at the
method level, using method call graphs, showing where indirect conflicts may occur. This type of exploratory design has
been used often in the visualization of indirect conflict tools, allowing developers a type of search area for their development
needs. Another indirect conflict tool, CASI~\cite{Servant:2010:CPI}, utilizes dependency slicing~\cite{Bajracharya:2009:SIS}
instead of method call graphs to provide an environment to see what areas of a project are being affected by a source code change.
Most of these tools have all shown to have the same common difficulties of scalability, false positives, and information overload,
which were explored in this paper. Our own tool
Impact!~\cite{Ell:2013} also suffered these same fates.

@@

For developer interest in regards to software modifications, Kim~\cite{Kim:2011:ESA} found that developers wanted to know whose
recent code changes semantically interfere with their own code changes, and whether their code is impacted by a code change.
Kim found that developers are concerned with interfaces of objects and when those interfaces change. Kim also identified the same 
issues towards information overload through false positives with developers noting ``I get a big laundry list... I see the email and I delete it''.
Kim's field study does however fall short in actually creating a resolution for indirect conflicts, or finding new concerns of
developers which are not already detected by compilation or other static analysis tools. 
For impact awareness, DeSouza et. al.~\cite{deSouza:2008:ESS} found that developers use their personal knowledge of the
code base to determine the impact of their code changes on fellow developers, teams, and projects. However, DeSouza does not
study which types of changes (types, frequencies, compounding factors) developers should concern themselves with more in terms of using their personal knowledge
to stop indirect conflicts. DeSouza also does not study formal mitigation strategies, or resolutions of indirect conflicts directly.

This paper is intended to fill the gap which has been left by aforementioned tools papers as well as the field study paper.
We will not only study why information overload, false positives, and scalability are such difficult problems to tackle
in indirect conflict tools, but we will also study how developers currently deal with indirect conflicts in practice through
their mitigation strategies, their largest concerns, and their ideas for future suggestions in regards to indirect conflicts.

@@

Our interview participants came from a large breadth of both open and closed source software development companies
and projects, using both  agile and waterfall based methodology, and from a wide spectrum of organizations, as shown in Table~\ref{tab:demo}.
Our participants were invited based on their direct involvement in the actual writing of software for
their respective companies or projects. These participants' software development experience had an average of 8 years of experience.
In addition to software development, some participants were also invited based on their experience with project management
at some capacity.

@@

Participants were invited to be interviewed by email and were sent a single reminder email one week
after the initial invitation if no response was made. We directly emailed 22 participants and  conducted
19 interviews. Interviews were conducted in person when possible, or over video chat when not possible, and recorded for audio content only.

@@

Interview participants first answered a number of demographic questions and were then asked to describe various software development experiences regarding our three research questions.
Specifically, ten semi-structured questions related to our research questions where ask and guided our interview:

@@

Our questionnaire respondents were different from our interviewees. We invited our questionnaire participants from a similar
breadth of open and closed source software development
companies and projects as the interviews participants with two main exceptions, the addition of GitHub users as well as Apache Software Foundation (Apache) developers.
The software organizations that remained the same between interview and questionnaire were: Mozilla, The GNOME Project, Microsoft Corporation,
Subnet Solutions, and Amazon. The GitHub
users were selected based on large amounts of development activity on GitHub and the Apache developers
were selected based on their software development contributions on specific projects known to be used heavily
utilized by other organizations and projects.

@@

Questionnaire participants were invited to participate in the questionnaire by email. We directly emailed 1300 participants and ended with 78 responses
giving a response rate of 6\%. We attribute the low response rate with: the questionnaires
were conducted during the months of July and August while many participants may be away from their regular positions.

The questionnaire we designed ~\footnote{http://thesegalgroup.org/people/jordan-ell/iced\_survey/}
was based on the insights we obtained from the round of interviews, and was intended to confirm some of these insights but also to broaden them to a larger sample size of developers who may have similar or different opinions from those already acquired from the interviews. The question topics asked in the questionnaire were:

@@

\subsection{Data Analysis}
To analyze both the interview and questionnaire data we used grounded theory techniques as described by Corbin and Strauss~\cite{Corbin:1998:SP}.
Grounded theory is a qualitative research methodology that utilizes \textit{theoretical sampling} and
\textit{open coding} to create a theory ``grounded'' in the empirical data (although this paper only uses grounded
theory techniques for analysis and not for a final theory). For an exploratory study such as
the ours, grounded theory is well suited because it involves starting from very broad and abstract type questions, and
making refinements along the way as the study progresses and hypotheses begin to take shape. Grounded theory involves
realigning the sampling criteria throughout the course of the study to ensure that participants are able to answer new
questions that have been formulated in regards to forming hypotheses. In our study being presented, data collected from
both interviews and questionnaires (when open ended questions were involved) was analyzed using open and axial coding (relating
code and concepts to each other).
In Section~\ref{sec:eval}, we give a brief evaluation of our studying using
3 criteria that are commonly used in evaluating grounded theory studies.

@@

Following our data collection and analysis, we re-interviewed 5 of our initial interview participants
in order to validate our findings. We confirmed our findings as to whether or not they resonate with
industry participants' opinions and experiences regarding indirect conflicts and as to their industrial
applicability. Due to limited time constraints of the interviewed participants, we could only re-interview
five participants. Those that were re-interviewed came from the range of 5-10 years of software development
experience. Re-interviewed participants were given our 3 research questions along with results and main
discussion points, and asked open ended questions regarding their opinions and experiences to validate our
findings. We also evaluated our grounded theory approach as per Corbin and Strauss'~\cite{Corbin:1998:SP}
list of criteria to evaluate quality and credibility. This evaluation can be seen in Section~\ref{sec:eval}

@@

We now present our results of both the interviews and questionnaires conducted in regards to our 3 research questions
outlined in Section~\ref{sec:intro}. We restate each research question, followed by our quantitative and qualitative
results from which we draw our discussion to be seen in Section~\ref{sec:disc}. In each subsection, quantitative data
given refers to interviews conducted unless explicitly said otherwise.

@@

A large contributing factor to the cause
of indirect conflicts comes from the changing of a software object's contract said 12 developers. Object contracts are, in a sense,
what a software object guarantees, meaning how the input, output, or how any aspect of the object is guaranteed
to work; made famous by Eiffel Software's~\footnote{http://www.eiffel.com/} ``Design by Contract''\texttrademark.
In light of object contracts, 14 developers gave examples of indirect conflicts they had experienced
which stemmed from not understanding the far reaching ramifications of a change being made to an object contract
towards the rest of the project. Of those 14, 3 dealt
with the changing of legacy code, with one developer saying ``legacy code does not change because developers
are afraid of the long range issues that may arise from that change''. Another developer, in regards to changing
object contracts stated ``there are no changes in the input or changes in the output, but the behavior is different''.
Developers also noted that the conflicts that do occur tend to be quite unique from each other and do not necessarily
follow common patterns.

In regards to object contract changes, 9 developers currently working with large scale database applications listed database
schemas as a large source of indirect conflicts while 5 developers that work on either software
library projects or are in test said that methods or functions were the root of their indirect conflict issues.
7 developers mentioned that indirect conflicts occur when a major update to an external project,
library, or service occurs with one developer noting ``their build never breaks, but it breaks ours''. Some
other notable indirect conflict artifacts were user interfaces in web development and full components in component
base game architecture.

@@

12 developers said that
when a project is in the early stages of development, indirect conflicts tend to occur far more frequently
than once a stable point is reached. Developers said ``At a stable point we decided we are not going to change
[this feature] anymore. We will only add new code instead of changing it.'' and ``the beginning of a project
changes a lot, especially with agile''. Questionnaire data also added ``indirect conflicts after a release
depend on how well the project was built at first'', ``[indirect conflicts] tend to slow down a bit after a
major release, unless the next release is a major rework.'', and ``[indirect conflicts have] spikes during
large revamps or the implementation of cross-cutting new features.'' which corroborate our interview results.
Questionnaire data also showed that
indirect conflicts are more likely to occur before the first major release rather than after at the daily
and weekly occurrence intervals as seen in Table~\ref{tab:often}.

@@

In terms of organizational structure, questionnaire data shows that as a project becomes larger and more
developers are added, even to the point that multiple teams are formed, indirect conflicts become more likely to
occur. However, indirect conflicts still occur at a lower number of developers as well with even 43\% of developers
saying they are like to occur in single developer projects. This can be seen in Table~\ref{tab:env}.

@@

In terms of preventative processes used for indirect conflicts, 3 major components were found.
First, design by contract is heavily used by interviewed developers as a means to avoid indirect conflicts or understand
when they are likely to occur. The use of design by contract here means that developers tend not to change an object's
contract when possible, and that an object's contract is used as a type of documentation towards awareness of the
software object. One developer stated that ``design by contract was invented to solve this problem and it does it
quite well'', while another noted that software object contracts do solve the problem in theory, but that doesn't
mean that problems don't still occur in practice.
Second, 21\% of developers mentioned that the add and deprecate
model is used to prevent indirect conflicts once
the project, feature, or component has reached a stable or mature point.
Add and deprecate meaning instead of editing code, the developer simply clones old code (if needed), and edits the clone
while slowly phasing out the out the old code in subsequent releases or ad needed. This allows other software to
use the older versions of software objects which remain unchanged, thus avoiding indirect conflicts.
Lastly, pure developer experience was mentioned with 7 developers mentioning that when planning code changes,
either a very experienced member of the project was involved in the planning and has duties to foresee any
indirect conflicts that may arise, or that developers must use their personal knowledge to predict where indirect
conflicts will occur while implementing.

@@

\bf{Being that developers believe ``good enough'' solutions exist for preventing and detecting indirect conflicts
developers would rather see improvement to debugging processes. Developers are looking for automated debugging tools,
and better source code analysis tool for impact management and supporting cross team and cross sub-project projects.}

@@

We now discuss the aforementioned
results in relation to the outstanding issues of tool based solutions regarding indirect conflicts which are information overload,
false positives, and scalability.
As per our results, three main discussion points have arisen. What developers really want, and don't want from indirect
conflict awareness tools, if indirect conflicts are better handled by extensively trying to prevent them or by simply
fixing them as they occur with less prevention, and a gap in analytical tools for software dependencies found in current
industry settings.

@@

As we have seen, indirect conflicts are found to be quite a serious problem that occur frequently,
sporadically, and differ greatly from case to case. These conditions pose large issues for the creation of generalizable theories
or tools in regards to indirect conflicts. These underlying complexities are the probable cause of disinterest of software developers
to proposed or implemented tools as discussed in Section~\ref{sec:related}. This inability to generalize is what we believe to be the
leading cause of information overload and false positives in the awareness system, causing developers to eventually ignore
information being presented to them, rendering the system useless.

\textbf{These false positives are caused by a difference in what
developers consider to be false positives versus what awareness literature considers they are.}

This disjoint, as will
be seen, is caused by developers only considering events which cause some action to be taken on their part, to be true positives,
where as current awareness understanding would state any event which is related to an individual's
work~\cite{Herbsleb06collaborationin}~\cite{Cataldo:2008:SCF}, to be true positives.
``You need a good understanding of what the code change is or else you will have a lot of false positives'' said one developer,
showing that not all changes around an object should be reported for awareness.

@@

\textbf{Developers don't care about what they don't need to account for}

``We would want a high probability of the [reported] change being an issue''.
This sense of unwanted or limited awareness is crucial to understand why generalized
awareness techniques of difficult to generalize problems, such as indirect conflicts through generalized changes to
classes~\cite{Sarma:2007:TSA}, or methods~\cite{Trainer:2005:BGT,Servant:2010:CPI}, often encounter difficulties of false
positives and information overload.
Developers simply do not want to know about events which effect them, but require no action on their part.

@@

We have seen through the results, that developers possess both tools and practices for the prevention,
detection, and resolution of indirect conflicts.

\textbf{Through unwanted awareness, and the use of developer time,
				developers tend to prefer working on real issues that have already occurred as opposed to preventing issues of the future which
				may never arise.}

This is neatly explained by the popular adage ``I work until something breaks'' taken by most developers. This mindset
is a clear example of developers taking a curative approach to software development opposed to a preventative approach.
Prevention here refers to taking precautionary steps to stop potential or some probability of issues, or indirect conflicts, form occurring in the first
place, while cure refers to fixing non probable (have happened) issues as they arise which includes not attempting, or putting little attempt, into preventing
them in the first place.

@@

\textbf{Prevention versus Cure} \textit{The need for continued study of the open question
``Which of prevention or cure is more effective for software development and indirect conflicts?''. This simple question will undoubtedly
produce extremely complex answers.}

With software developers focusing their current efforts on curative measures, for
indirect conflicts, suggests
that while it may not be the most effective, it may be the easiest road for developers to take. This may be the path of
least resistance, but it may not be optimal. Software engineering as a whole should strive to answer this question or
provide more insight into possibilities, as its answer may determine many future actions taken by the research community.

@@

\textbf{Awareness Theory to Model Real World} \textit{the need to further characterize the
mismatch between awareness approaches and tools as developed in the research community and awareness needs as perceived
by industry. We should push our awareness theories to model real development
environments where interruptions, time management, and full development life cycles are often large factors.}

@@

\textbf{Give Developers What They Want} \textit{Developers have their own notions of ``good''
that something is for their productivity. We should pay attention to what they want, but also be mindful that what
developers want may not always be good for them.}

The direct implication this research has on tool creation is that of current adoption among developers. As was stated,
developers are more keen to invest their time at the detection and cure / resolution stages of indirect conflicts. That being
said, focus at these two stages for tool developer will lead to larger adoption among developers.
It should be noted here that detection must come with an almost zero rate of false positives as
current tools (unit and integration and user testing), while they may not have 100\% recall, have almost 100\% precision.

@@

{\bfseries Fit.} ``Do the findings fit/resonate with the professional for whom the research was intended and the participants?'' This
criterion is used to verify the correctness of our finding and to ensure they resonate and fit with participant opinion. It is also
required that the results are generalizable to all participants but not so much as to dilute meaning. To ensure fit, during interviews
after participants gave their own opinions on a topic, we presented them with previous participant opinions and asked them to comment
on and potentially agree with what the majority has been on the topic. Often the developers own opinions already matched those of
the majority before them and did not necessarily have to directly verify it themselves.

As added insurance, we conducted 5 post results interviews with developers to once again confirm our results, and discussions. These
procedures can be seen in Section~\ref{sec:meth}.

@@

Indirect conflicts are a significant issue with real world development, however, many proposed techniques and tools to mitigate
losses in this realm have been unsuccessful in attracting major support from developers. Based on our qualitative study involving
19 interviewed developers from 12 organizations as well as 78 developers completing questionnaires, we have provided characterization of indirect conflicts,
current developer work flow surrounding indirect conflicts, and what direction any future tools should follow in order to aid developers
in their work flow with indirect conflicts.

>>>>

% Paper title
% Can use linebreaks \\ within to get better formatting as desired
\title{State of The Practice in Indirect Conflicts: Causes, Mitigation Strategies, and Future Directions}

@@

As Software Configuration Management (SCM) has grown, the maturity and norm of parallel
development has become the standard development process instead of the exception. With this parallel development
comes the need for larger awareness among developers to have ``an understanding of the activities of others
which provides a context for one's own activities''~\cite{Dourish:1992:ACS}. This added awareness
mitigates some downsides of parallel development which include the cost of conflict prevention and resolution. However,
empirical evidence shows that despite of recent research has proposed mitigation strategies, these losses continue to appear quite frequently and can prove to be a significant
and time-consuming chore for developers~\cite{Perry:2001:PCL, Sarma:2009:TIV}.

@@

While indirect conflict tools have shown potential from studies of developers, some of the same problems continue
to arise throughout most, if not all tools. The most prevalent issue is that of false positives and information
overload, which causes tools to eventually be
ignored~\cite{Sarma:2007:TSA, Servant:2010:CPI}. A second primary issue is that of code dependency identification and
tracking. Many different dependencies have been proposed and used in indirect conflict tools such as method
invocation~\cite{Trainer:2005:BGT}, and class signatures~\cite{Sarma:2007:TSA} with varying success; however, the
identification of failure inducing changes, other than those which are already identifiable by other means such
as compilers or unit tests, to these dependencies still remains an issue.
Some tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA} have also leveraged the notion of socio-technical congruence ~\cite{Cataldo:2006:ICR} to help developers pursue dependencies through social means.
However, the same aforementioned problems still arise. Evaluations of these tools or approaches have largely been small in scale, and involved a limited number of industrial participants or student evaluators. As a result, the problems of scalability, information overload and false positives in tackling indirect conflicts remains an outstanding research problem.

In this paper we report from a field study of software practitioners that sought to characterize the state-of-the-practice in indirect conflicts, particularly their causes and mitigation strategies. Our intention has been to uncover the root causes of information overload, false positives, and
scalability issues, in order to better understand why indirect conflict tools and methods fail to fully achieve
their intended success. Three research questions that guided our study:

@@

To answer these questions, we interviewed 19 developers from across 12 commercial and open source projects, sought confirmation of the themes we identified in our interviews through a questionnaire with 78
developers, as well as 5 follow-up interviews. Our findings indicate that:
\begin{itemize}
	\item Indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding of such changes
	\item Developers tend to prefer to use detection and resolution processes or tools over those of prevention
	\item Developers do not want awareness mechanisms which provide non actionable results
	\item There exists a gap in software evolution analytical tools from the reliance on static analysis resulting in missed
				context of indirect conflicts
\end{itemize}

@@

Many indirect conflict tools have been developed, tested, and published. Sarma et al.~\cite{Sarma:2007:TSA} created Palantir,
which can both detect potential indirect conflicts, at the class signature level, and alert developers to these conflicts.
Palantir represented one of the first serious attempts at aiding developers towards indirect conflicts. Holmes et
al.~\cite{Holmes:2010:CAR} take it one step further with their tool YooHoo, by detecting fine grained source code changes,
such as method return type changes, and create a taxonomy for different types of changes and their proneness to cause
indirect conflicts. This tool and its taxonomy had a severely reduced false positive rate, however, the true positives
detected may already be detectable by current tools such as compilers and unit testing. The tool Ariadne~\cite{Trainer:2005:BGT}
creates an environment where developers can see how source code changes will affect other areas of a project at the
method level, using method call graphs, showing where indirect conflicts may occur. This type of exploratory design has
been used often in the visualization of indirect conflict tools, allowing developers a type of search area for their development
needs. Another indirect conflict tool, CASI~\cite{Servant:2010:CPI}, utilizes dependency slicing~\cite{Bajracharya:2009:SIS}
instead of method call graphs to provide an environment to see what areas of a project are being affected by a source code change.
The evaluations of these tools indicate that they have the same common difficulties of scalability, false positives, and information overload. Our own tool
Impact!~\cite{Ell:2013} also suffered these same fates.

@@

For developer interest in regards to software modifications, Kim~\cite{Kim:2011:ESA} found that developers wanted to know whose
recent code changes semantically interfere with their own code changes, and whether their code is impacted by a code change.
Kim found that developers are concerned with interfaces of objects and when those interfaces change. Kim also identified the same
issues towards information overload through false positives with developers noting ``I get a big laundry list... I see the email and I delete it''.
Kim's field study does however fall short in actually creating a resolution for indirect conflicts, or finding new concerns of
developers which are not already detected by compilation or other static analysis tools.
In studying awareness of change impact, deSouza et. al.~\cite{deSouza:2008:ESS} found that developers use their personal knowledge of the
code base to determine the impact of their code changes on fellow developers, teams, and projects. However, their work does not
study which types of changes (types, frequencies, compounding factors) developers encounter in their work, nor provides concrete information on formal mitigation strategies or resolutions of indirect conflicts.

The insights of our study intends to fill the gap. In our field study we seek to understand not only  why information overload, false positives, and scalability are such difficult problems to tackle
in indirect conflict tools, but also how developers currently deal with indirect conflicts in practice through
their mitigation strategies, what are their major concerns as well as suggestions for future research and tools to tackle these concerns.

@@

Our interview participants came from a wide range of both open and closed source software development companies
and projects, using both  agile and waterfall based methodologies, and from a wide spectrum of organizations, as shown in Table~\ref{tab:demo}.
The participants had software development experience in average of 8 years, as well as some project management experience.

@@

Participants were invited based on their software engineering and project management experience. We recruited participants through an email message to a number of software organizations in our working business network, and followed up in a single reminder email one week
after the initial invitation if no response was made. We directly emailed 22 participants and  conducted
19 interviews. Interviews were conducted in person when possible, or over video chat when not possible, and recorded for audio content only.

@@

Interview participants first answered a number of demographic questions and were then asked to describe various software development experiences and practices.
Specifically, ten semi-structured questions related to our research questions where ask and guided our interview:

@@

Our questionnaire respondents were different from our interviewees. We recruited the respondents to our questionnaire from a similar
breadth of open and closed source software development
companies and projects as the interviews participants with two main exceptions. The GitHub users were recruited from the Google BigQuery database; the Apache respondents were recruited from the list of tops contributors to the Apache Software Foundation.
The software organizations that remained the same between interview and questionnaire were: Mozilla, The GNOME Project, Microsoft Corporation,
Subnet Solutions, and Amazon. The GitHub
users were selected based on large amounts of development activity on GitHub and the Apache developers
were selected based on their software development contributions on specific projects known to be used heavily
utilized by other organizations and projects.

@@

Questionnaire participants were invited to participate in the questionnaire by email. We directly emailed 1300 participants and obtained 78 responses
giving a response rate of 6\%. We attribute the low response rate with: the questionnaires
were conducted during the months of July and August while many participants may be away from their regular positions.

The questionnaire we designed ~\footnote{http://thesegalgroup.org/projects/indirect-conflicts/}
was based on the insights we obtained from the round of interviews, and was intended to confirm these insights but also to broaden them to a larger sample size of developers who may have similar or different opinions from those already acquired from the interviews. The question topics asked in the questionnaire were:

@@

\subsection{Data Analysis}
To analyze both the interview and questionnaire data we used grounded theory techniques as described by Corbin and Strauss~\cite{Corbin:1998:SP}.
Grounded theory is a qualitative research methodology that utilizes \textit{theoretical sampling} and
\textit{open coding} to create a theory ``grounded'' in the empirical data. In this study we used grounded
theory techniques for analysis and not to develop a theory per say. For an exploratory study such as ours, grounded theory is well suited because it involves starting from very broad and abstract type questions, and
making refinements along the way as the study progresses and hypotheses begin to take shape. Grounded theory involves
realigning the sampling criteria throughout the course of the study to ensure that participants are able to answer new
questions that have been formulated in regards to forming hypotheses. In our study being presented, data collected from
both interviews and questionnaires (when open ended questions were involved) was analyzed using open and axial coding (relating
code and concepts to each other).
In Section~\ref{sec:eval}, we give a brief evaluation of our studying using
3 criteria that are commonly used in evaluating grounded theory studies.

@@

Following our data collection and analysis, we re-interviewed 5 of our initial interview participants
in order to validate our findings. We confirmed our findings as to whether or not they resonate with
industry participants' opinions and experiences regarding indirect conflicts and as to their industrial
applicability. Due to limited time constraints of the interviewed participants, we could only re-interview
five participants. Those that were re-interviewed came from the range of 5-10 years of software development
experience. Re-interviewed participants were given our 3 research questions along with results and main
discussion points, and asked open ended questions regarding their opinions and experiences to validate our
findings. We also evaluated our grounded theory approach as per Corbin and Strauss'~\cite{Corbin:1998:SP}
list of criteria to evaluate quality and credibility, described in Section~\ref{sec:eval}.

@@

In this section we present our results of both the interviews and questionnaires conducted in regards to our 3 research questions
outlined in Section~\ref{sec:intro}. For each research question we summarize the quantitative and qualitative
evidence we obtained. In each subsection, quantitative data
given refers to interviews conducted unless explicitly said otherwise.

@@

We found that a large contributing factor to the cause
of indirect conflicts comes from the changing of a software object's contract (reported by 12 developers directly). Object contracts, or ``Design by Contract''\texttrademark, a concept made well-known through the work on Eiffel Software's~\footnote{http://www.eiffel.com/} are, in a sense,
what a software object guarantees, meaning how the input, output, or how any aspect of the object is guaranteed.
In our interviews, 14 developers gave examples of indirect conflicts they had experienced
which stemmed from not understanding the far reaching ramifications of a change being made to an object contract
towards the rest of the project. Of those 14, 3 dealt
with the changing of legacy code, with one developer saying ``legacy code does not change because developers
are afraid of the long range issues that may arise from that change''. Another developer, in regards to changing
object contracts stated ``there are no changes in the input or changes in the output, but the behavior is different''.
Developers also noted that the conflicts that do occur tend to be quite unique from each other and do not necessarily
follow common patterns.

In regards to object contract changes, 9 developers currently working with large scale database applications listed database
schemas as a large source of indirect conflicts while 5 developers that work on either software
library projects or are in test said that methods or functions were the root of their indirect conflict issues.
7 developers mentioned that indirect conflicts occur when a major update to an external project,
library, or service occurs with one developer noting that ``their build never breaks, but it breaks ours''. Some
other notable indirect conflict artifacts were user interfaces in web development and full components in component
base game architecture.

@@

12 developers reported that
when a project is in the early stages of development, indirect conflicts tend to occur far more frequently
than once a stable point is reached. For example, ``At a stable point we decided we are not going to change
[this feature] anymore. We will only add new code instead of changing it.'' and ``the beginning of a project
changes a lot, especially with agile''. From the survey respondents we also learned that ``indirect conflicts after a release
depend on how well the project was built at first'', ``[indirect conflicts] tend to slow down a bit after a
major release, unless the next release is a major rework.'', and ``[indirect conflicts have] spikes during
large revamps or the implementation of cross-cutting new features.'' which corroborate our interview results.
Questionnaire data also showed that
indirect conflicts are more likely to occur before the first major release rather than after at the daily
and weekly occurrence intervals as seen in Table~\ref{tab:often}.

@@

In terms of organizational structure, questionnaire data shows that as a project becomes larger and more
developers are added, even to the point that multiple teams are formed, indirect conflicts become more likely to
occur. However, as shown in Table~\ref{tab:env}, indirect conflicts still occur in small teams, with  43\% of respondents indicating
 they are likely to occur in single developer projects.

@@

Our analysis indicates indicates three major preventative processes used for indirect conflicts.
First, design by contract is heavily used by interviewed developers as a means to avoid indirect conflicts or to understand
when they are likely to occur. The use of design by contract here means that developers tend not to change an object's
contract when possible, and that an object's contract is used as a type of documentation towards awareness of the
software object. One developer stated that ``design by contract was invented to solve this problem and it does it
quite well'', while another noted that software object contracts do solve the problem in theory, but that doesn't
mean that problems don't still occur in practice.
Second, 21\% of developers mentioned that the add and deprecate
model is used to prevent indirect conflicts once
the project, feature, or component has reached a stable or mature point.
Add and deprecate meaning instead of editing code, the developer simply clones old code (if needed), and edits the clone
while slowly phasing out the out the old code in subsequent releases or ad needed. This allows other software to
use the older versions of software objects which remain unchanged, thus avoiding indirect conflicts.
Lastly, pure developer experience was mentioned with 7 developers mentioning that when planning code changes,
either a very experienced member of the project was involved in the planning and has duties to foresee any
indirect conflicts that may arise, or that developers must use their personal knowledge to predict where indirect
conflicts will occur while implementing.

@@

\bf{Because developers believe ``good enough'' solutions exist for preventing and detecting indirect conflicts, we identified a strong preference for improvements to debugging processes. Developers prefer automated debugging tools
and better source code analysis tool for impact management and supporting cross team and cross sub-project projects.}

@@

We now discuss these results in relation to the outstanding issues of tool-based solutions regarding indirect conflicts and which are information overload,
false positives, and scalability.
As per our findings, three main discussion points emerge: (1) What developers really want and don't want from indirect
conflict awareness tools, (2) Whether indirect conflicts are better handled by extensively trying to prevent them or by simply
fixing them as they occur with less prevention, and (3) A gap in analytical tools for software dependencies found in current
industry settings.

@@

According to our study participants, indirect conflicts represent a serious problem that occur frequently,
sporadically, and differ greatly from case to case. These conditions pose large issues for the creation of generalizable theories
or tools in regards to indirect conflicts. These underlying complexities are the probable cause of disinterest of software developers
to proposed or implemented tools as discussed in Section~\ref{sec:related}. This inability to generalize is what we believe to be the
leading cause of information overload and false positives  awareness systems, causing developers to eventually ignore
information being presented to them, rendering the system useless.

\textbf{These false positives are caused by a difference in what
developers consider to be false positives versus what awareness literature considers they are.} This disjoint, we argue, is caused by developers only considering events which cause some action to be taken on their part, to be true positives,
where as current awareness understanding would state any event which is related to an individual's
work~\cite{Herbsleb06collaborationin}~\cite{Cataldo:2008:SCF}, to be true positives.
``You need a good understanding of what the code change is or else you will have a lot of false positives'' said one developer,
showing that not all changes around an object should be reported for awareness.

@@

\textbf{Developers don't care about what they don't need to account for}. ``We would want a high probability of the [reported] change being an issue''.
This sense of unwanted or limited awareness is crucial to understand why generalized
awareness techniques of difficult to generalize problems, such as indirect conflicts through generalized changes to
classes~\cite{Sarma:2007:TSA}, or methods~\cite{Trainer:2005:BGT,Servant:2010:CPI}, often encounter difficulties of false
positives and information overload.
Developers simply do not want to know about events which effect them, but require no action on their part.

@@

Our findings suggest that developers possess both tools and practices for the prevention,
detection, and resolution of indirect conflicts.

\textbf{Through unwanted awareness, and the use of developer time,
				developers tend to prefer working on real issues that have already occurred as opposed to preventing issues of the future which
				may never arise.} Related to the popular adage ``I work until something breaks'' taken by most developers, this mindset
is a clear example of developers taking a curative approach to software development opposed to a preventative approach.
Prevention here refers to taking precautionary steps to stop potential or some probability of issues, or indirect conflicts, form occurring in the first
place, while cure refers to fixing non probable (have happened) issues as they arise which includes not attempting, or putting little attempt, into preventing
them in the first place.

@@

\textbf{Prevention versus Cure} \textit{There is a need for continued study of the open question
``Which of prevention or cure is more effective for software development and indirect conflicts?''. We believe that this simple question will undoubtedly
produce extremely complex answers. Developers are currently favoring cure, but we should be mindful that what developers want
may not always be the best choice.}

The fact that software developers focus their current efforts on curative measures, for
indirect conflicts, suggests
that while it may not be the most effective, it may be the easiest road for developers to take. This may be the path of
least resistance, but it may not be optimal. Software engineering as a whole should strive to answer this question or
provide more insight into possibilities, as its answer may determine many future actions taken by the research community.

@@

\textbf{Theories of Awareness to Model Real World} \textit{There is a need to further characterize the
mismatch between awareness approaches and tools as developed in the research community and awareness needs as perceived
by industry. Theories of awareness should consider modeling real development
environments where interruptions, time management, and full development life cycles are often large factors.}

@@

\textbf{Give Developers What They Want} \textit{Developers have their own notions of how ``good''
something is for their productivity. We should pay attention to what they want, in order to shape our
models and better address their needs.}

The direct implication this research has on tool development is that of current adoption among developers. As was stated,
developers are more keen to invest their time at the detection and cure / resolution stages of indirect conflicts. That being
said, focus at these two stages for tool developer will lead to larger adoption among developers.
It should be noted here that detection must come with an almost zero rate of false positives as
current tools (unit and integration and user testing), while they may not have 100\% recall, have almost 100\% precision.

@@

{\bfseries Fit.} ``Do the findings fit/resonate with the professional for whom the research was intended and the participants?'' This
criterion is used to verify the correctness of our finding and to ensure they resonate and fit with participant opinion. It is also
required that the results are generalizable to all participants but not so much as to dilute meaning. To ensure fit, during interviews
after participants gave their own opinions on a topic, we presented them with previous participant opinions and asked them to comment
on and potentially agree with what the majority has been on the topic. Often the developers' own opinions already matched those of
the majority before them and did not necessarily have to directly verify it themselves.

As added insurance, we conducted 5 confirmatory interviews after we analyzed our data and drafted themes in our findings. We describe this
procedure in Section~\ref{sec:meth}.

@@

Indirect conflicts are a significant issue with real world development, however, many of our research proposed techniques and tools have been unsuccessful in attracting major support from developers. Based on our qualitative study involving
19 interviewed developers from 12 organizations as well as 78 questionnaires respondents, we sought to characterize indirect conflicts,
current developer work flows surrounding indirect conflicts, and frame directions for future tool development that support developers
in their tackling of indirect conflicts.
